---
title: "Project2ajh4539"
author: "Amanda Hassell"
date: "4/29/2021"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(echo = TRUE, eval = TRUE,fig.align="center",warning=FALSE,message=FALSE,fig.width=8, fig.height=5, linewidth=60)
options(tibble.width = 100,width = 100)
library(tidyverse)
```

Introduction:

- **0. (5 pts)** Introduce your dataset and each of your variables (or just your main variables if you have lots) in a paragraph. What are they measuring? How many observations?

```{r}
library(tidyverse)
install.packages('fivethirtyeightdata', repos = 'https://fivethirtyeightdata.github.io/drat/', type = 'source')
library(fivethirtyeightdata)
library(fivethirtyeight)

view(bechdel)

bechdel2<- na.omit(bechdel)
```
This dataset (bechdel) observes individual movies from 1990-2013. The dataset is based on the Bechdel test, which observes the presence of women in Hollywood movies. With this test, movies either pass or fail (as represetnted by the 'binary' variable) based on how many women are present and named in the film, whether the women talk to eachother in the film, and whether the women are talking about something other than men. It is important to note that this is not a definitive way of determinign gender equity in Hollywood movies, but it is the best test available that is aimed at doing so. The clean_test variable has 5 categories based on how a movie scores in the Bechdel test. 'Ok' means the movie passed the test, 'dubious' means it is unclear as to whether the movie passed the test, 'men' means there are more than 2 named women who talk to each other but they only talk about men, 'notalk' means there are more than 2 named women but they dont talk to eachtoher, and 'nowomen' means that there are less than 2 named women in the movie. There are also numeric variables including 'budget_2013', 'domgross_2013', and 'intgross_2013', which observe the movies' budget, domestic gross profit, and international gross profit, respectively. The numbers are adjusted for inflation as of the year 2013. The goal of this project is to draw conclusions about the funding of movies taking into account their representation of women, and to determine whether movies that exclude women make more, less, or equal profits as do movies that do not. After omitting all observations containing containing NAs, this dataset includes 1600 observations.Each observation/row is a discrete movie title.

MANOVA:
```{r}
library(mvtnorm); library(ggExtra)
df<-rmvnorm(1000,mean=c(0,0),sigma=matrix(c(1,.5,.5,1),ncol=2,byrow=T))
df<-data.frame(df)%>%rename(Y1=X1,Y2=X2)
p<-ggplot(df, aes(Y1,Y2))+geom_point(alpha=.5)+geom_density_2d(h=2)+coord_fixed()
ggMarginal(p,type="density",xparams = list(bw=.5), yparams=list(bw=.5))

ggplot(bechdel2, aes(x = budget_2013, y = domgross_2013)) + geom_point(alpha = .5) + geom_density_2d(h=2) + coord_fixed() + facet_wrap(~clean_test)
ggplot(bechdel2, aes(x = budget_2013, y = intgross_2013)) + geom_point(alpha = .5) + geom_density_2d(h=2) + coord_fixed() + facet_wrap(~clean_test)
ggplot(bechdel2, aes(x = domgross_2013, y = intgross_2013)) + geom_point(alpha = .5) + geom_density_2d(h=2) + coord_fixed() + facet_wrap(~clean_test)

library(rstatix)
group <- bechdel2$clean_test
DVs <- bechdel2 %>% select(budget_2013,domgross_2013,intgross_2013)
#Test multivariate normality for each group (null: normality met)
sapply(split(DVs,group), mshapiro_test)
lapply(split(DVs,group), cov) #separate covariance matrices for each group
box_m(DVs, group)

bechdelmanova<- manova(cbind(budget_2013, domgross_2013, intgross_2013) ~ clean_test, data = bechdel2)
summary(bechdelmanova)
summary.aov(bechdelmanova)

bechdel2%>%group_by(clean_test)%>%summarize(mean(budget_2013),mean(domgross_2013), mean(intgross_2013))

pairwise.t.test(bechdel2$budget_2013, bechdel2$clean_test, p.adj="none")
pairwise.t.test(bechdel2$domgross_2013, bechdel2$clean_test, p.adj="none")
pairwise.t.test(bechdel2$intgross_2013, bechdel2$clean_test, p.adj="none")

#total 30 t tests + 3 ANOVA + 1 MANOVA = 34 total tests 
# α = 0.05/34 = 0.0015
```
total tests = 30 t tests + 3 ANOVA + 1 MANOVA = 34 total tests 
bonferroni α = 0.05/34 = 0.0015
The adjusted significance level (bonferroni correction) is 0.0015. This significance level is used in all subsequent analyses. 

The overall MANOVA was significant (Pillai trace = 0.039395, pseudo F(12,4785) = 5.3059, p < 0.0001). Univariate ANOVAs for each dependent variable were conducted as follow-up tests to the MANOVA, using the Bonferroni method for controlling Type I error rates for multiple comparisons. Only one univariate ANOVA (for budget_2013) showed a significant mean difference across levels of the clean_test categorical variable (F(4,1595) = 12.816, p < 0.0001). 

Post hoc analysis was performed conducting pairwise comparisons (t-tests) to determine which clean_test results differed in budget_2013. Movies who passed the bechdel test (ok) were found to differ significantly from movies in which women did not talk to each other (no talk) and movies which did had less than 2 women (no women) in terms of budget. 

MANOVA has multiple assumptions, most of which appear to have been met. This includes random samples and independent observations. Since each movie is it's own row or observation, this assumption is met. Another assumption is Multivariate normality of DVs and linear relationships among DVs, which appears to be true. Within group co-variance matrices are homogeneous. There are no extreme univariate or multivariate outliers and DVs don't appear to be too correlated. 


Randomization Test:
```{r}
ggplot(bechdel2,aes(domgross_2013,fill=binary))+geom_histogram(bins=100)+
facet_wrap(~binary,ncol=2)+theme(legend.position="none")

# Ho = mean domestic gross profit (adjusted for inflation) is the same for movies who passed vs. failed the Bechdel test
# Ha = mean domestic gross profit (adjusted for inflation) is different for movies who passed vs. failed the Bechdel test

bechdel2%>% group_by(binary) %>% summarize(means=mean(domgross_2013)) %>% summarize(diff(means))

rand_dist<-vector()
for(i in 1:5000){
  new<-data.frame(domgross_2013=sample(bechdel2$domgross_2013),binary=bechdel2$binary)
  rand_dist[i]<-mean(new[new$binary=="PASS",]$domgross_2013)- mean(new[new$binary=="FAIL",]$domgross_2013)}

mean(rand_dist>18901999 | rand_dist < -18901999) #p-value

{hist(rand_dist,main="",ylab=""); abline(v = c(-14768357, 14768357),col="red")}

```
We failed to reject the null hypothesis. This means that there is not a significant difference in mean domestic gross profit between movies that passed vs failed the Bechdel test. The practical application to this is that we should have no reason to believe that excluding women in Hollywood movies leads to increased (or decreased) profits from those movies here in the United States.  

Linear Regression Model:
```{r}
library(tidyverse)

#linear regression model 
bechdel2$budget_2013_c <- bechdel2$budget_2013 - mean(bechdel2$budget_2013)
mean(bechdel2$budget_2013)
bechdelfit<- lm(domgross_2013~budget_2013_c*binary, data = bechdel2)
summary(bechdelfit) #get R-squared values to explain variation

#plot of regression
ggplot(bechdel2, aes(budget_2013_c,domgross_2013, color = binary)) + geom_smooth(method = "lm", se = F, fullrange = T) + geom_point()+geom_vline(xintercept=0,lty=2)

#check assumptions
resids<-bechdelfit$residuals
fitvals<-bechdelfit$fitted.values
ggplot()+geom_point(aes(fitvals,resids))+geom_hline(yintercept=0, color='red')
library(sandwich); library(lmtest)
bptest(bechdelfit) #formally confirms heteroskedasticity
ggplot() +geom_histogram(aes(resids), bins = 100)

#robust SEs
summary(bechdelfit)$coef[,1:2]
coeftest(bechdelfit, vcov = vcovHC(bechdelfit))[,1:2]
```
A movie with a $0 budget does not have a resonable application, so this variable was mean centered (budget_2013_c). The intercept coefficient (8.154e+07) corresponds to the mean/predicted domestic gross profit for a movie that failed the Bechdel test with an average budget (57,456,564 dollars). For movies that have an average budget and pass the Bechdel test, the domestic gross profit is 7.289e+06 times higher than those who don't pass the Bechdel test. The estimated slope of budget on domestic gross profit for movies that failed the Bechdel test is 1.086e+00. The coefficent for budget_2013_c:binaryPASS is the difference in slopes between budget and movies that passed the Bechdel test. For every one unit increase in budget, predicted domestic gross proifit goes up 1.515e-01 for this group. 

The proportion of variation in the outcome that this model explains is 0.4147 (multiple R-squared). After adjusting for having multiple explanatory variables in the model, this proportion comes to 0.4126 (adjusted R-squared). This adjusted estimate is more conservative than the multiple R-squared. 

Based on the fanning pattern from the graph, the data violates the homoskedasticity assumption but linearity looks okay. Normality also looks okay based on the second graph. 

Since the data was heteroskedastic, robust standard errors were corrected and the results were similar. The coefficient estimates remained the same as the original model. However, the corrected standard errors are slightly larger and therefore should be used for the sake of being more conservative. 

Bootstrapping:
```{r}
samp_distn<-replicate(5000, {
  boot_dat <- sample_frac(bechdel2, replace=T) 
  fitboot <- lm(domgross_2013~budget_2013_c+binary, data=boot_dat)
  coef(fitboot) })

samp_distn %>% t %>% as.data.frame %>% summarize_all(sd)

boot_dat <- sample_frac(bechdel2, replace=T) 
fitboot <- lm(domgross_2013~budget_2013_c+binary, data=boot_dat)
summary(fitboot)
```
The bootstrapped standard errors for the intercept, budget_2013 mean centered variable, and binaryPASS condition are 2585312, 0.05989816, and 3951322, respectively. P-values and standard errors remain very similar (compared to both the original SEs and the robust SEs). 


Logistic Regression Model:
```{r}
library(lmtest)
data<-bechdel2%>%mutate(y=ifelse(binary=="FAIL",1,0))
data$binary<-factor(data$binary,levels=c("FAIL","PASS"))
#1 = fail bechdel test
#0 = pass bechdel test

fitlog<-glm(binary~budget_2013+intgross_2013, data=data, family="binomial")
coeftest(fitlog)

coef(fitlog)%>%data.frame #log odds scale coefficients
coef(fitlog)%>%exp%>%round(5)%>%data.frame #odds scale coefficients

odds2prob<-function(odds){odds/(1+odds)} #function to convert odds to probs
odds2prob(1.24732) #probabiltiy of failing Bechdel test when budget and intgross are zero
odds2prob(1)
 
probs<-predict(fitlog,type="response") #get predicted probs from the model
#using .5 as the threshold for predicting failing the bechdel test
table(predict=as.numeric(probs>.5),truth=data$y)%>%addmargins

#Accuracy = (true positives + true negatives)/ total population  
331+337 #668

#sensitivity(TPR): probability of predicting a failure if a movie really did fail the test. 
331/856 #0.3866822

#specificity (TNR): probability of predicting a pass for movies that really passed the test
337/744 #0.452957

#Precision (PPV): the proportion classified as failing the test that actually did
331/738 #0.4485095

#denstiy plot
data$logit<- predict(fitlog, type = "link")
data %>% mutate(y = as.factor(y)) %>% ggplot()+geom_density(aes(logit, color = y, fill = y), alpha = 0.4)+ geom_vline(xintercept = 0) +xlab("logit(logodds)")

#ROC curve
data$probs <-predict(fitlog,type="response")
library(plotROC)
ROCplot<-ggplot(data)+geom_roc(aes(d=y,m=probs), n.cuts=0)
ROCplot
calc_auc(ROCplot)
```
The coefficient for budget_2013 indicates that increasing by 1 dollar in budget multiplies odds by a factor of about 1.000. When converted to probability, this means that there is about a 50% chance of failing the Bechdel test when international gross profit is zero. The result is the same for the intgross_2013 coefficient and correlates to a 50% chance of failing the bechdel test when budget is zero. The coefficient for the intercept indicates that when budget and international gross profit are zero, there is a 0.5550255 probability of failing the Bechdel test.

Accuracy, Sensitivity (TPR), Specificity (TNR), Precision (PPV), and AUC of the model are 668, 0.3866822, 0.452957, 0.4485095, and 0.3955743, as outlined and explained above. The AUC for the ROC curve is 0.3955743. 

Logistic Regression of ALL Variables:
```{r}
#in-sample logistic regression
fitlog2<-glm(binary~budget_2013+intgross_2013+domgross_2013, data=data, family="binomial")
coeftest(fitlog2)

coef(fitlog2)%>%data.frame #log odds scale coefficients
coef(fitlog2)%>%exp%>%data.frame #odds scale coefficients 

probs2<-predict(fitlog2,type="response") #get predicted probs from the model
#using .5 as the threshold for predicting failing the bechdel test
table(predict=as.numeric(probs2>.5),truth=data$y)%>%addmargins

odds2prob(1.25713)

#Accuracy = (true positives + true negatives)/ total population  
334+342 #676

#sensitivity(TPR): probability of predicting a failure if a movie really did fail the test. 
334/856 #0.3901869

#specificity (TNR): probability of predicting a pass for movies that really passed the test
342/744 #0.4596774

#Precision (PPV): the proportion classified as failing the test that actually did
334/736 #0.4538043

data$probs2 <-predict(fitlog2,type="response")
ROCplot2<-ggplot(data)+geom_roc(aes(d=y,m=probs2), n.cuts=0)
ROCplot2
calc_auc(ROCplot2)
```
The coefficient for budget_2013 indicates that increasing by 1 dollar in budget multiplies odds by a factor of about 1.000. When converted to probability, this means that there is about a 50% chance of failing the Bechdel test when international gross profit and domestic gross profit are zero. The result is the same for the intgross_2013 coefficient and correlates to a 50% chance of failing the bechdel test when budget and domgestic gross profit are zero. Again, the same result occurrs for domgross_2013, in which there is a 50% chance of failing the test when budget and international gross profit are both zero. Finally, the coefficient for the intercept indicates that when budget and international gross profit are zero, there is a 0.5569595 probability of failing the Bechdel test.
Accuracy, Sensitivity (TPR), Specificity (TNR), Precision (PPV), and AUC of the model are 676, 0.3901869, 0.4596774, 0.4538043, and 0.3957156, as outlined and explained above. The AUC  is based on all possible cutoffs/traeoffs and essentially summarizes sensitivity and specificity, so since the AUC is low here, this means that the model is not predicting very well. 

```{r}
#class_diag function
class_diag<-function(probs,truth){
  
  tab<-table(factor(probs>.5,levels=c("FALSE","TRUE")),truth)
  acc=sum(diag(tab))/sum(tab)
  sens=tab[2,2]/colSums(tab)[2]
  spec=tab[1,1]/colSums(tab)[1]
  ppv=tab[2,2]/rowSums(tab)[2]

  if(is.numeric(truth)==FALSE & is.logical(truth)==FALSE) truth<-as.numeric(truth)-1
  
  #CALCULATE EXACT AUC
  ord<-order(probs, decreasing=TRUE)
  probs <- probs[ord]; truth <- truth[ord]
  
  TPR=cumsum(truth)/max(1,sum(truth)) 
  FPR=cumsum(!truth)/max(1,sum(!truth))
  
  dup<-c(probs[-1]>=probs[-length(probs)], FALSE)
  TPR<-c(0,TPR[!dup],1); FPR<-c(0,FPR[!dup],1)
  
  n <- length(TPR)
  auc<- sum( ((TPR[-1]+TPR[-n])/2) * (FPR[-1]-FPR[-n]) )

  data.frame(acc,sens,spec,ppv,auc)
}

#10 fold CV
k=10

cvdata<-data[sample(nrow(data)),] 
folds<-cut(seq(1:nrow(data)),breaks=k,labels=F) 
diags<-NULL

for(i in 1:k){
train<-cvdata[folds!=i,]
test<-cvdata[folds==i,]
truth<-test$binary

fitlog2<-glm(binary~budget_2013+intgross_2013+domgross_2013, data=data, family="binomial")

probs3<-predict(fitlog2,newdata = test,type="response")

diags<-rbind(diags,class_diag(probs3,truth))
}

summarize_all(diags,mean) 
```
The accuracy, sensitivity, specificity, precision, and AUC values for the 10-fold cross validation model are 0.5775, 0.5410074, 0.6101294, 0.5470705, and 0.6048802, respectively. This AUC is much higher than that of the previous logistic regression in-sample model (0.3957156). The AUC is based on all possible cutoffs/traeoffs and essentially summarizes sensitivity and specificity, so since the AUC is higher here, this means that the model is predicting better than the in-sample model. 

```{r}
#LASSO
library(glmnet)

y<-as.matrix(data$binary)
x<-model.matrix(binary~ budget_2013+ domgross_2013+ intgross_2013,data=data)[,-1]
head(x)
x<- scale(x)

cv<-cv.glmnet(x,y,family="binomial")

{plot(cv$glmnet.fit, "lambda", label = TRUE); abline(v = log(cv$lambda.1se)); abline(v = log(cv$lambda.min), lty=2)}

lasso<-glmnet(x,y,family="binomial",lambda=cv$lambda.1se)
coef(lasso)

```
After conduction a LASSO, the only variable which was retained is the budget_2013 variable. 

```{r}
#10 fold CV with only LASSO chosen variable (budget_2013)
k=10

cvdata<-data[sample(nrow(data)),] 
folds<-cut(seq(1:nrow(data)),breaks=k,labels=F) 
diags<-NULL

for(i in 1:k){
train<-cvdata[folds!=i,]
test<-cvdata[folds==i,]
truth<-test$binary

fitlog2<-glm(binary~budget_2013, data=data, family="binomial")

probs3<-predict(fitlog2,newdata = test,type="response")

diags<-rbind(diags,class_diag(probs3,truth))
}

summarize_all(diags,mean) 
```
The new AUC value, using only the budget_2013 variable, is 0.6026587. This value is very similar to that of the original 10-fold cross validation, and is also much higher than that of the in-sample model from the original logistic regression. This means that this model is predicting better than the original logistic regression model. 
